Cultural Transmission Can Explain the Evolution of Cooperative Behavior

### 2020-07-19 ms/main.tex
fix tables; remove wA wB

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage{longtable}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=CadetBlue,citecolor=black,linkcolor=black]{hyperref}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage{authblk}

@@

Consider a very large population whose members are characterized by their phenotype $\phi$, which can be of two types, $\phi=A$ for cooperators or $\phi=B$ for defectors, with corresponding fitness values $w_A$ and $w_B$, which depend on the frequency of the phenotypes (see below).
An offspring inherits its phenotype from its parent via vertical transmission with probability $v$ or from a random individual in the parental population via oblique transmission with probability $(1-v)$. 
Following~\citet{ram2018evolution}, given that the parent phenotype is $\phi$ and assuming uni-parental inheritance, % yr: cite Zefferman Behav Ecol 2016
the conditional probability that the phenotype $\phi'$ of the offspring is $A$ is

@@

Individuals interact according to a prisoner's dilemma.
Specifically, individuals interact in pairs; a cooperator suffers a fitness cost $0<c<1$, and its partner gains a fitness benefit $b$, where we assume $b>c$. The following payoff matrix shows the fitness of an individual with phenotype $\phi_1$ when interacting with a partner of phenotype $\phi_2$ ($b>c>0$):

\begin{table}[h]\label{table:prisoner_payoff}
\centering
\begin{tabular}{lll}
           & $\phi_2=A$ & $\phi_2=B$ \\
$\phi_1=A$ & $b-c$      & $-c$       \\
$\phi_1=B$ & $b$        & $0$       
\end{tabular}
\end{table}

@@

%Therefore, the fitness values of the two phenotypes, $w_A$ and $w_B$ are
%
%\begin{equation}\label{eq:fitness}
%\begin{aligned}
%w_A(\hat{p}) & = \hat{p}(1+b-c) + (1-\hat{p})(1-c) = 1 - c + \hat{p} b \\
%w_B(\hat{p}) & = \hat{p}(1+b) + (1-\hat{p}) = 1 + \hat{p} b,
%\end{aligned}
%\end{equation}
%and the population mean fitness is therefore
%$\bar{w} = \hat{p} w_A + (1-\hat{p}) w_B = 1 + \hat{p}(b-c)$.

Horizontal cultural transmission occurs between peers. 
It may occur between social partners with probability $\alpha$, or between a random pair with probability $1-\alpha$ (see~\textbf{\autoref{fig:horizontal}}).
Horizontal transmission is not always successful, as one peer may reject the other's phenotype. The probability for successful transmission of phenotypes $A$ and $B$ are $T_A$ and $T_B$, respectively.\\
\autoref{table:1} contains the probability of first interactor to be A following interaction by interaction type.
Therefore, the frequency $p'$ of phenotype $A$ among adults in the next generation, after horizontal transmission, is 
\begin{equation}\label{eq:nextgen_adults}
\begin{aligned}
p'
& = \hat{p}^2 [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A] \;,
\end{aligned}
\end{equation}
which simplifies to
\begin{equation}\label{eq:nextgen_adults_slimpify}
p' = \hat{p}^2(T_B-T_A) + \hat{p}(1+T_A-T_B) .
\end{equation}

The frequency of $A$ among parents (after selection) follows a similar dynamic, but also includes the effect of natural selection, and is therefore
\begin{equation}\label{eq:nextgen_parents}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) (1-c) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} (1+b) [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A] \;,
\end{aligned}
\end{equation}
where fitness values are taken from \autoref{table:1}, and 
\begin{equation} \label{eq:mean_fitness}
\bar{w} = \hat{p} w_A + (1-\hat{p}) w_B = 1 + \hat{p}(b-c)
\end{equation}  %MWF: Define w_A, w_B.
is the population mean fitness.
Eq~\ref{eq:nextgen_parents} can be simplified to
\begin{equation}\label{eq:nextgen_parents_simplified}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) \big(1-(1-\hat{p})(1-\alpha)T_B)\big) \\
& + \hat{p}(1-\hat{p}) (1-c) \big(\hat{p}(1-\alpha)T_B+1-T_B\big) \\
& + (1-\hat{p})\hat{p} (1+b) \big(\hat{p}(1-\alpha) + \alpha\big) T_A \\
& + (1-\hat{p})^2 \hat{p} (1-\alpha) T_A \;.
\end{aligned}
\end{equation}

\begin{table}[H]
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Interaction Type} & \textbf{Frequency}    & \textbf{Fitness}      & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Probability of first interactor\\  to be A %MWF: What is A?
 following interaction\end{tabular}}}                                                                                                     \\ \hline
\multicolumn{1}{|l|}{}    & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \textbf{\begin{tabular}[c]{@{}c@{}}Horizontal transmission\\  from partner, \\ probability $\alpha$\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Horizontal transmission \\ from population, \\ probability $1-\alpha$\end{tabular}} \\ \hline
$AxA$                     & $\hat{p}^2$           & $1+b-c$               & $1$                                                                                                               & $\hat{p}+(1-\hat{p})(1-T_B)$                                                                                           \\ \hline
$AxB$                     & $\hat{p}(1-\hat{p})$  & $1-c$                 & $1-T_B$                                                                                                           & $\hat{p}+(1-\hat{p})(1-T_B)$                                                                                           \\ \hline
$BxA$                     & $\hat{p}(1-\hat{p})$  & $1+b$                 & $T_A$                                                                                                             & $\hat{p}T_A$                                                                                                           \\ \hline
$BxB$                     & $(1-\hat{p})^2$       & $1$                   & $0$                                                                                                               & $\hat{p}T_A$                                                                                                           \\ \hline
\end{tabular}
\caption{Fitness and probability of the first interactor to be A following interaction by interaction type} %MWF: What is A?
\label{table:1}
\end{table}

\begin{figure}[b]
  \centering
  \includegraphics[scale=1]{figure1.pdf}
  \caption{\textbf{Cultural horizontal transmission.} Transmission occurs between interacting partners with probability $\alpha$ (left) or between two random peers with probability $1-\alpha$.}
  \label{fig:horizontal}
\end{figure}

>>>>

\usepackage{amssymb,amsmath,amsthm}
\usepackage{commath}
\usepackage[hyphens]{url}
\usepackage[dvipsnames]{xcolor}
\usepackage[unicode=true,colorlinks=true,urlcolor=CadetBlue,citecolor=black,linkcolor=black]{hyperref}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage{authblk}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{booktabs}

@@

Consider a very large population whose members are characterized by their phenotype $\phi$, which can be of two types, $\phi=A$ for cooperators or $\phi=B$ for defectors.
An offspring inherits its phenotype from its parent via vertical transmission with probability $v$ or from a random individual in the parental population via oblique transmission with probability $(1-v)$. 
Following~\citet{ram2018evolution}, given that the parent phenotype is $\phi$ and assuming uni-parental inheritance, % yr: cite Zefferman Behav Ecol 2016
the conditional probability that the phenotype $\phi'$ of the offspring is $A$ is

@@

Individuals interact according to a prisoner's dilemma.
Specifically, individuals interact in pairs; a cooperator suffers a fitness cost $0<c<1$, and its partner gains a fitness benefit $b$, where we assume $b>c$. \autoref{table:prisoner_payoff} shows the payoff matrix, i.e. the fitness, of an individual with phenotype $\phi_1$ when interacting with a partner of phenotype $\phi_2$ (we assume $b>c>0$):

% Table: payoff matrix

\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
           & $\phi_2=A$ & $\phi_2=B$ \\ \cmidrule(r){1-3}
$\phi_1=A$ & $1+b-c$ & $1-c$ \\
$\phi_1=B$ & $1+b$   & $1$
\\ \bottomrule
\end{tabular}
\caption{\textbf{Payoff matrix for prisoner's dilemma.}
The fitness of phenotype $\phi_1$ when interacting with phenotype $\phi_2$. $A$ is a cooperative phenotype, $B$ is a defector phenotype, $b$ is the benefit gained by an individual interacting with a cooperator, and $c$ is the cost of cooperation. $b>c>0$.
}
\label{table:prisoner_payoff}
\end{table}

@@

Horizontal cultural transmission occurs between peers. 
It may occur between social partners with probability $\alpha$, or between a random pair with probability $1-\alpha$ (see~\textbf{\autoref{fig:horizontal}}).
Horizontal transmission is not always successful, as one peer may reject the other's phenotype. The probability for successful transmission of phenotypes $A$ and $B$ are $T_A$ and $T_B$, respectively.\\
\autoref{table:interactions} contains the probability of first interactor to be A following interaction by interaction type.
Therefore, the frequency $p'$ of phenotype $A$ among adults in the next generation, after horizontal transmission, is 
\begin{equation}\label{eq:nextgen_adults}
\begin{aligned}
p'
& = \hat{p}^2 [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A] \;,
\end{aligned}
\end{equation}
which simplifies to
\begin{equation}\label{eq:nextgen_adults_slimpify}
p' = \hat{p}^2(T_B-T_A) + \hat{p}(1+T_A-T_B) .
\end{equation}

The frequency of $A$ among parents (after selection) follows a similar dynamic, but also includes the effect of natural selection, and is therefore
\begin{equation}\label{eq:nextgen_parents}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) (1-c) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} (1+b) [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A] \;,
\end{aligned}
\end{equation}
where fitness values are taken from \autoref{table:interactions}, and 
\begin{equation} \label{eq:mean_fitness}
\bar{w} =  1 + \hat{p}(b-c)
\end{equation}
is the population mean fitness.
Eq~\ref{eq:nextgen_parents} can be simplified to
\begin{equation}\label{eq:nextgen_parents_simplified}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) \big(1-(1-\hat{p})(1-\alpha)T_B)\big) \\
& + \hat{p}(1-\hat{p}) (1-c) \big(\hat{p}(1-\alpha)T_B+1-T_B\big) \\
& + (1-\hat{p})\hat{p} (1+b) \big(\hat{p}(1-\alpha) + \alpha\big) T_A \\
& + (1-\hat{p})^2 \hat{p} (1-\alpha) T_A \;.
\end{aligned}
\end{equation}

% Table: interactions
\begin{table}[]
\begin{tabular}{@{}llllll@{}}
\toprule
\multirow{2}{*}{Phenotype $\phi_1$} &
  \multirow{2}{*}{Phenotype $\phi_2$} &
  \multirow{2}{*}{Frequency} &
  \multirow{2}{*}{Fitness of $\phi_1$} &
  \multicolumn{2}{l}{$P(\phi_1=A)$ via horizontal transmission:} \\ \cmidrule(l){5-6} 
    &     &                      &         & from partner, $\alpha$ & from population, $(1-\alpha)$ \\ \cmidrule(r){1-6}
$A$ & $A$ & $\hat{p}^2$          & $1+b-c$ & 1                      & $\hat{p}+(1-\hat{p})(1-T_B)$  \\
$A$ & $B$ & $\hat{p}(1-\hat{p})$ & $1-c$   & $1-T_B$                & $\hat{p}+(1-\hat{p})(1-T_B)$  \\
$B$ & $A$ & $\hat{p}(1-\hat{p})$ & $1+b$   & $T_A$                  & $\hat{p} T_A$                 \\
$B$ & $B$ & $(1-\hat{p})^2$      & $1$     & $0$                    & $\hat{p} T_A$                 \\ \bottomrule
\end{tabular}
\caption{\textbf{Interaction frequency, fitness, and transmission probabilities.}}
\label{table:interactions}
\end{table}

% Figure: model illustration
\begin{figure}[b]
  \centering
  \includegraphics[scale=1]{figure1.pdf}
  \caption{\textbf{Cultural horizontal transmission.} Transmission occurs between interacting partners with probability $\alpha$ (left) or between two random peers with probability $1-\alpha$.}
  \label{fig:horizontal}
\end{figure}
